---
title: "Compare Two Variances - F-test"
author: "YOSEPH"
date: "2018년 11월 28일"
output: html_document
---

### 1. **F-Test**

#### (1) Goal
To be used to assess **whether the variances of two populations (A and B) are equal**.
		
#### (2) When to use ?
To compare two variances is useful in several cases, including:

- When you want to perform a **two samples t-test** to check the equality of the variances of the two samples
- When you want to compare the variability of a new measurement method to an old one.  
Does the new method reduce the variability of the measure?


### 2. **Research questions** and **statistical hypotheses**

#### (1) Typical research questions are:
a. whether the variance of group A (σ^2^~A~) is **equal** to the variance of group B (σ^2^~B~)?
b. whether the variance of group A (σ^2^~A~) is **less** than the variance of group B (σ^2^~B~)?
c. whether the variance of group A (σ^2^~A~) is **greather** than the variance of group B (σ^2^~B~)?

#### (2) In statistics, we can define the corresponding **null hypothesis (H~0~)** as follow:
a. H~0~ : σ^2^~A~ = σ^2^~B~ (**two-tailed** test)
b. H~0~ : σ^2^~A~ > σ^2^~B~ (**one-tailed** test)
c. H~0~ : σ^2^~A~ < σ^2^~B~ (**one-tailed** test)
  
#### (3) the corresponding **alternative hypotheses (H~a~)** are as follow:
a. H~a~ : σ^2^~A~ ≠ σ^2^~B~ (**different**)
b. H~a~ : σ^2^~A~ ≤ σ^2^~B~ (**greater**)
c. H~a~ : σ^2^~A~ ≥ σ^2^~B~ (**less**)


### 3. **Formula** of `F-test`  

The test statistic can be obtained by computing the ratio of the **two variances**, S^2^~A~ and S^2^~B~:

> F = S^2^~A~ / S^2^~B~

The **degrees of freedom** are n~A~−1 (for the numerator) and n~B~−1 (for the denominator).

- **Note 1** : that the **more** this ratio **deviates from 1**, the **stronger** the evidence for **unequal** population variances.
- **Note 2** : that the F-test requires the **two samples** to be **normally distributed**.


### 4. Compute `F-test`

	* R function
		  `var.test()` can be used to compare **two variances** as follow:

	* Two Methods

		  * Method 1 : `var.test(values ~ groups, data, alternative = "two.sided")`
		  * Method 2 : `var.test(x, y, alternative = "two.sided")`

		    x,y : numeric vectors
		    
		    alternative: the alternative hypothesis. 
					           allowed value is one of “two.sided” (**default**), “greater” or “less”.
<p/>

### 5. Import and check your data

	* Here, we’ll use the built-in R data set named ToothGrowth:
		
```{r code1, message=FALSE, warning=FALSE, include=TRUE, echo=TRUE}

# Store the data in the variable my_data
data(ToothGrowth)

my_data <- ToothGrowth
```

	* To have an idea of what the data look like, we start by displaying a random sample of 10 rows using the function sample_n{dplyr} :

```{r code2, message=FALSE, warning=FALSE, include=TRUE, echo=TRUE}

require(dplyr)

sample_n(my_data, 10)
```

	* We want to test the equality of variances between the two groups OJ and VC in the column “supp”
<p/>

### 6. Preliminary test to check `F-test` assumptions
	
	* `F-test` is very sensitive to departure from the normal assumption.
		  You need to check whether the data is normally distributed 
		  before using the `F-test`

	* `Shapiro-Wilk` test can be used to test whether the normal assumption holds. 
		  It’s also possible to use `Q-Q plot` (quantile-quantile plot) to graphically 
		  evaluate the normality of a variable.

		  `Q-Q plot` draws the correlation between a given sample and the normal distribution

	* If there is doubt about normality, the better choice is to use `Levene’s test` or `Fligner-Killeen test`, 
	  which are less sensitive to departure from normal assumption
<p/>

### 7. Compute `F-test`  

```{r code3, message=TRUE, warning=FALSE, include=TRUE, echo=TRUE}

# F-test
( res.ftest <- var.test(len ~ supp, data = my_data) )
```
<p/>

### 8. Interpretation of the result

	The p-value of `F-test` is p = 0.2331433 which is greater than the significance level 0.05. 
	In conclusion, there is no significant difference between the two variances.
<p/>

### 9. Access to the values returned by `var.test()` function

	The function `var.test()` returns a list containing the following components:

	  - statistic : the value of the F test statistic
	  - parameter : the degrees of the freedom of the F distribution of the test statistic
	  - p.value 	: the p-value of the test
	  - conf.int 	: a confidence interval for the ratio of the population variances
	  - estimate 	: the ratio of the sample variances
<p/>

	The format of the R code to use for getting these values is as follow:
	
```{r code4, message=TRUE, warning=FALSE, include=TRUE, echo=TRUE}

# ratio of variances
res.ftest$estimate

# p-value of the test
res.ftest$p.value
```