---
title: "5-2. Correlation Matrix"
output: html_document
---

## Correlation matrix : to analyze, format and visualize a correlation matrix

#
    There are different methods for correlation analysis : 
    1. Pearson correlation test (parametric)
    2. Spearman rank-based correlation analysis (non-parametric)
    3. Kendall rank-based correlation analysis (non-parametric)


(@) ###What is correlation matrix?

#
    A correlation matrix, which is used to investigate the dependence between multiple variables at the same time. 
    The result is a table containing the correlation coefficients between each variable and the others
    
    
(@) ###Compute correlation matrix

+ ####R functions

    - The R function `cor{stats}` can be used to compute a correlation matrix. 
    - A simplified format of the function is :

```{r code1, eval=FALSE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
cor(x, method = c("pearson", "kendall", "spearman"))
```
#
    1. x      : numeric matrix or a data frame.
    2. method : indicates the correlation coefficient to be computed. 
    3. The default is pearson correlation coefficient which measures the linear dependence between two variables. 
    4. kendall and spearman correlation methods are non-parametric rank-based correlation test.
    
+ ####Missing Values (NA)

    - If data contain missing values, 
      use the following R code to handle missing values by case-wise deletion.

```{r code2, eval=FALSE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
cor(x, method = "pearson", use = "complete.obs")
```

(@) ###Import your data into R

Here, we’ll use a data derived from the built-in R data set `mtcars{datasets}` as an example:

```{r code3, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# Load data
data("mtcars")

# select the next columns in a mtcars data set:
#   mpg, disp, hp, drat, wt, qsec
my_data <- mtcars[, c(1,3,4,5,6,7)]

# print the random sampling observations using sample_n{dplyr}
library(dplyr, quietly = TRUE)
rsd <- sample_n(tbl = my_data, size = 10, replace = FALSE)

library(knitr)
kable(x = rsd, caption = 'A random sampling data of a my_data')
```

(@) ###Compute correlation matrix

In the following table, correlations coefficients between the possible pairs of variables are shown.

#
    *Note* if data contain missing values, 
           use the following R code to handle missing values by case-wise deletion.
    
    cor(my_data, use = "complete.obs")
    
#
    *Note* Unfortunately, 
           the function `cor{stats}` returns only the correlation coefficients between variables. 
           In the next section, we will use Hmisc R package to calculate the correlation p-values.

```{r ode4, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
( res <- cor(my_data) )
round(res, 2)
```

(@) ###Correlation matrix with significance levels (p-value)

The function `rcorr{Hmisc}` can be used to compute the significance levels for pearson and spearman correlations.  
It returns both the correlation coefficients and the p-value of the correlation for all possible pairs of columns in the data table.

+ Simplified format:

#
    rcorr(x, type = c("pearson", "spearman"))
    
    1. x should be a matrix. 
    2. The correlation type can be either pearson or spearman.
    
+ Install `{Hmisc}` package:
```{r code4, eval=FALSE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
install.packages("Hmisc")
```

+ Use `rcorr{Hmisc}` function:
```{r code5, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
library("Hmisc")

( res2 <- rcorr(as.matrix(my_data), type='pearson') )
```

#
    The output of the function `rcorr{Hmisc}` is a list containing the following elements : 
      - r : the correlation matrix 
      - n : the matrix of the number of observations used in analyzing each pair of variables 
      - P : the p-values corresponding to the significance levels of correlations.

If you want to extract the p-values or the correlation coefficients from the output, use this:
```{r code6, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# Extract the correlation coefficients
res2$r

# Extract p-values
res2$P
```

(@) ###A simple function to format the correlation matrix

#
    This section provides a simple function for formatting a correlation matrix 
    into a table with 4 columns containing :

      - Column 1 : row names (variable 1 for the correlation test)
      - Column 2 : column names (variable 2 for the correlation test)
      - Column 3 : the correlation coefficients
      - Column 4 : the p-values of the correlations
      
The custom function below can be used :
```{r code7, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# ++++++++++++++++++++++++++++
# flattenCorrMatrix
# ++++++++++++++++++++++++++++
# cormat : matrix of the correlation coefficients
# pmat : matrix of the correlation p-values

flattenCorrMatrix <- function(cormat, pmat) {
  # upper.tri{base} : Lower and Upper Triangular Part of a Matrix
  #   - return class : matrix
  #   - including logical values in all each cell of the returned matrix
  ( ut <- upper.tri(cormat) )
  
  # create a data frame
  #   - row{base} : Row Indexes
  #   - col{base} : Column Indexes
  #   - rownames{base} : Row Names
  #   - colnames{base} : Column Names
  data.frame(
      row = rownames(cormat)[row(cormat)[ut]],
      column = rownames(cormat)[col(cormat)[ut]],
      cor  =(cormat)[ut],
      p = pmat[ut]
  )
} # flattenCorrMatrix
```

Example of usage :
```{r code8, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
library(Hmisc, quietly = TRUE)

res2 <- rcorr( as.matrix( mtcars[ ,1:7] ) )
flattenCorrMatrix(res2$r, res2$P)
```

(@) ###Visualize correlation matrix

There are different ways for visualizing a correlation matrix :

+ `symnum{stats}` function : Symbolic Number Coding
    - Symbolically encode a given numeric or logical vector or array.
    - Particularly useful for visualization of structured matrices,
        - e.g., correlation, sparse, or logical ones.
+ `corrplot{corrplot}` function : A visualization of a correlation matrix
    - to plot a correlogram
+ scatter plots
+ heatmap

(@) ###Use `symnum{stats}` function: Symbolic number coding

The R function `symnum{stats}` replaces correlation coefficients by symbols according to the level of the correlation. 

It takes the correlation matrix as an argument :

+ Simplified format:
```{r code9, eval=FALSE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
symnum(
    x,
    cutpoints = c(0.3, 0.6, 0.8, 0.9, 0.95),
    symbols = c(" ", ".", ",", "+", "*", "B"),
    abbr.colnames = TRUE
)
```

#
    1. x              : the correlation matrix to visualize
    2. cutpoints      : correlation coefficient cutpoints
       The correlation coefficients between 0 and 0.3 are replaced by a space (" “),
       correlation coefficients between 0.3 and 0.6 are replace by ”.“, etc …
    3. symbols        : the symbols to use
    4. abbr.colnames  : logical value
                        If TRUE, colnames are abbreviated
                        
+ Example of usage:
```{r code10, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=TRUE}
# res : correlation matrix
symnum(res, abbr.colnames = FALSE)
```

#
    As indicated in the legend, 
      the correlation coefficients between 0 and 0.3 are replaced by a space (" “); 
      correlation coefficients between 0.3 and 0.6 are replace by”.“; 
      etc …

(@) ###Use `corrplot{corrplot}` function: Draw a correlogram

The function `corrplot{corrplot}`, in the package of the same name, creates a graphical display of a correlation matrix, highlighting the most correlated variables in a data table.

In this plot, correlation coefficients are colored according to the value. 
Correlation matrix can be also reordered according to the degree of association between variables.

+ Install corrplot:
```{r code11, eval=FALSE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
install.packages("corrplot", quiet = TRUE)
```

+ Use `corrplot{corrplot}` to create a correlogram:

The function `corrplot{corrplot}` takes the correlation matrix as the first argument.  
The second argument ( `type=“upper”` ) is used to display only the upper triangular of the correlation matrix.

```{r code12, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
library(corrplot)

# corrplot{corrplot} : A visualization of a correlation matrix.
#   res : correlation matrix
#   type : type of matrix to plot, c('full', 'lower', 'upper')
#   tl.col : text label color
#   tl.srt : text label string rotation in degrees
corrplot(res, type = "upper", order = "hclust", tl.col = "black", tl.srt = 45)
```

#
    The correlation matrix is reordered according to the correlation coefficient using “hclust” method.
      - tl.col (for text label color) and tl.srt (for text label string rotation) are used to change 
        text colors and rotations
      - Possible values for the argument type are : “upper”, “lower”, “full”

#
    1. Positive correlations are displayed in blue and negative correlations in red color. 
    2. Color intensity and the size of the circle are proportional to the correlation coefficients. 
    3. In the right side of the correlogram, 
       the legend color shows the correlation coefficients and the corresponding colors.

It’s also possible to combine correlogram with the significance test. 

We’ll use the result `res.cor2` generated in the previous section with `rcorr{Hmisc}` function :

#
    In the following plot, correlations with p-value > 0.01 or p-value > 0.05 are considered as insignificant. 
    In this case the correlation coefficient values are leaved blank or crosses are added.

```{r code13, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# Insignificant correlation are crossed
corrplot(res2$r, type="upper", order="hclust", p.mat = res2$P, sig.level = 0.01, insig = "blank")

# Insignificant correlations are leaved blank
corrplot(res2$r, type="lower", order="hclust", p.mat = res2$P, sig.level = 0.05, insig = "blank")
```


(@) ###Use `chart.Correlation{PerformanceAnalytics}` : Draw scatter plots

The function `chart.Correlation{PerformanceAnalytics}`, can be used to display a chart of a correlation matrix.

+ Install PerformanceAnalytics :
```{r code14, eval=FALSE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
install.packages("PerformanceAnalytics", quiet = TRUE)
```

+ Use `chart.Correlation{PerformanceAnalytics}` :
```{r code15, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
library("PerformanceAnalytics", quietly = TRUE)

my_data <- mtcars[, c(1,3,4,5,6,7)]
chart.Correlation(my_data, histogram=TRUE, pch=19)
```

#
    In the above plot:

    The distribution of each variable is shown on the diagonal.
    On the bottom of the diagonal : the bivariate scatter plots with a fitted line are displayed
    On the top of the diagonal    : the value of the correlation plus the significance level as stars
    Each significance level is associated to a symbol : p-values(0, 0.001, 0.01, 0.05, 0.1, 1) <=> symbols(“***”, “**”, “*”, “.”, " “)
    
(@) ###Use `heatmap{stats}`
```{r code16, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# Get some colors

# colorRampPalette{grDevices} : Color interpolation
#   These functions return functions that interpolate a set of given colors to create new color palettes (like topo.colors) and color ramps, functions that map the interval [0, 1] to colors
col<- colorRampPalette( c("blue", "white", "red") )(20)

# heatmap{stats} : draw a Heat Map
#   - symm : logical indicating if x should be treated symmetrically; 
#            can only be true when x is a square matrix.
heatmap(x = res, col = col, symm = TRUE)
```

#
    - x     : the correlation matrix to be plotted
    - col   : color palettes
    - symm  : logical indicating if x should be treated symmetrically; 
              can only be true when x is a square matrix.
