---
title: "5-3. Visualize correlation matrix using correlogram"
output: html_document
---

(@) ###Introduction

+ This article describes how to plot a correlogram.  
+ **Correlogram** is a graph of correlation matrix.  
+ It is very useful to highlight the **most correlated variables** in a data table.   
+ In this plot, **correlation coefficients** is colored according to the value.  
+ Correlation matrix can be also **reordered** according to the **degree of association** between variables.  
+ The R `{corrplot}` package is used here.

(@) ###Install R corrplot package

+ `{corrplot}` package is required to execute the R code in this article.

```{r code1, eval=FALSE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
install.packages("corrplot", quiet = TRUE)
```

(@) ###Data for correlation analysis

+ The `mtcars{datasets}` data is used to compute correlation matrix.

```{r code2, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
library(dplyr, quietly = TRUE)
( rs <- sample_n(tbl = mtcars, size = 10, replace = FALSE) )

library(knitr, quietly = TRUE)
kable(x = rs, caption = 'The random sampling of mtcars data set')
```

(@) ###Computing correlation matrix

```{r code3, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
( M <- cor( mtcars ) )
head( round(M, 2) )
```

(@) ###Correlogram : Visualizing the correlation matrix

R `corrplot{corrplot}` function is used to plot the graph of the correlation matrix.

+ The simplified format of the function is :
```{r code4, eval=FALSE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
library(corrplot, quietly = TRUE)

corrplot( corr, method="circle" )
```

#
    - corr    : The correlation matrix to visualize. 
                To visualize a general matrix, please use is.corr=FALSE
    - method  : The visualization method : “circle”, “color”, “number”, etc


(@) ###Visualization methods

+ Seven different visualization methods can be used : “circle”, “square”, “ellipse”, “number”, “shade”, “color”, “pie”.
```{r code5, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
library(corrplot, quietly = TRUE)

corrplot(M, method="circle")
corrplot(M, method="square")
corrplot(M, method="ellipse")
corrplot(M, method="shade")
corrplot(M, method="color")
corrplot(M, method="pie")
```

#
    1. Positive correlations are displayed in blue and negative correlations in red color. 
    2. Color intensity and the size of the circle are proportional to the correlation coefficients.

+ Display the correlation coefficient :
```{r code6, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
corrplot(M, method="number")
```

(@) ###Types of correlogram layout

+ There are three types of layout :
    - `“full”` (default)  : display full correlation matrix
    - `“upper”`           : display upper triangular of the correlation matrix
    - `“lower”`           : display lower triangular of the correlation matrix

```{r code7, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
corrplot(M, type="upper")
corrplot(M, type="lower")
```

(@) ###Reordering the correlation matrix

+ The correlation matrix can be reordered according to the correlation coefficient.  
+ This is important to identify the hidden structure and pattern in the matrix.  
+ `“hclust”` for hierarchical clustering order is used in the following examples.  

```{r code8, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# correlogram with "hclust" reordering
corrplot(M, type="upper", order="hclust")

# Using different color spectrum
col<- colorRampPalette(c("red", "white", "blue"))(20)
corrplot(M, type="upper", order="hclust", col=col)

# Change background color to lightblue
corrplot(M, type="upper", order="hclust", col=c("black", "white"), bg="lightblue")
```

(@) ###Changing the color of the correlogram

+ As shown in the above section, the color of the `correlogram` can be customized.
+ `RcolorBrewer` palette of colors are used in the R script below :

```{r code9, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
library(RColorBrewer, quietly = TRUE)

# brewer.pal{RColorBrewer} : ColorBrewer palettes
#                            Creates nice looking color palettes especially for thematic maps
corrplot(M, type="upper", order="hclust", col=brewer.pal(n=8, name="RdBu" ))
corrplot(M, type="upper", order="hclust", col=brewer.pal(n=8, name="RdYlBu"))
corrplot(M, type="upper", order="hclust", col=brewer.pal(n=8, name="PuOr"))
```


(@) ###Changing the color and the rotation of text labels

+ `tl.col` (for text label color) and `tl.srt` (for text label string rotation) are used to change text colors and rotations.

```{r code10, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
corrplot(M, type="upper", order="hclust", tl.col="blue", tl.srt=45)
```


(@) ###Combining correlogram with the significance test

+ ####Computing the p-value of correlations
    - To compute the matrix of p-value, a custom R function is used :
```{r code11, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# mat : is a matrix of data
# ... : further arguments to pass to the native R cor.test{stats} function
cor.mtest <- function(mat, ...) {
    mat <- as.matrix(mat)
    n <- ncol(mat)
    p.mat<- matrix(NA, n, n)
    diag(p.mat) <- 0
    
    for (i in 1:(n - 1)) {
        for (j in (i + 1):n) {
            tmp <- cor.test(mat[, i], mat[, j], ...)
            p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
        }
    }
    
    colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
    p.mat
}

# matrix of the p-value of the correlation
p.mat <- cor.mtest(mtcars)
head(p.mat[, 1:5])
```
    
+ ####Add significance level to the correlogram

#
    In the below figure,
      - correlations with p-value > 0.01 are considered as insignificant. 
      - In this case, the correlation coefficient values are leaved blank or crosses are added.

```{r code12, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
# Specialized the insignificant value according to the significant level
corrplot(M, type="upper", order="hclust", p.mat = p.mat, sig.level = 0.01)

# Leave blank on no significant coefficient
corrplot(M, type="upper", order="hclust", p.mat = p.mat, sig.level = 0.01, insig = "blank")
```


(@) ###Customize the correlogram
```{r code13, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))

corrplot(
      M,
      method="color",
      col=col(200),
      type="upper",
      order="hclust", 
      # Add coefficient of correlation
      addCoef.col = "black",
      #Text label color and rotation
      tl.col="red",
      tl.srt=45,
      # Combine with significance
      p.mat = p.mat,
      sig.level = 0.01,
      insig = "blank",
      # hide correlation coefficient on the principal diagonal
      diag=FALSE 
  )
```


(@) ###Conclusions

+ Use **`corrplot{corrplot}`** function to plot an elegant graph of a correlation matrix