---
title: '5-5. Correlation matrix : An R function to do all you need'
output: html_document
---

Correlation matrix analysis is very useful to study dependences or associations between variables.  
This article provides a custom R function, `rquery.cormat()`, for calculating and visualizing easily acorrelation matrix.  

The result is a list containing, the correlation coefficient tables and the p-values of the correlations.  
In the result, the variables are reordered according to the level of the correlation which can help to quickly identify the most associated variables.  

A graph is also generated to visualize the correlation matrix using a correlogram or a heatmap.

(@) ###Prerequisites

+ The `rquery.cormat()` function requires the installation of `{corrplot}` package. 
+ Before proceeding, install it using he following R code :

```{r code1, eval=FALSE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
install.packages("corrplot")
```

+ To use the `rquery.cormat()` function, you can source it as follow :
```{r code2, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
source("scripts/rquery_cormat.r")
```

+ A simplified format of the function is :
```{r code1_1, eval=FALSE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
rquery.cormat(
              x,
              type=c('lower', 'upper', 'full', 'flatten'),
              graph=TRUE,
              graphType=c("correlogram", "heatmap"),
              col=NULL,
              ...
        )
```

#
    Description of the arguments:

      (1) x         : matrix of data values
      (2) type      : Possible values are “lower” (default), “upper”, “full” or “flatten”. 
                      It displays the lower or upper triangular of the matrix, full or flatten matrix.
      (3) graph     : if TRUE,
                      a correlogram or heatmap is generated to visualize the correlation matrix.
      (4) graphType : Type of graphs. 
                      Possible values are “correlogram” or “heatmap”.
      (5) col       : colors to use for the correlogram or the heatmap.
      (6) …         : Further arguments to be passed to cor() or cor.test() function.

(@) ###Example of data

+ The `mtcars{datasets}` data is used in the following examples :
```{r code3, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
mydata <- mtcars[, c(1,3,4,5,6,7)]

library(dplyr, quietly = TRUE)
sample_n(tbl = mydata, size = 10, replace = FALSE)
```


(@) ###Computing the correlation matrix

```{r code4, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
rquery.cormat(mydata)
```

#
    The result of rquery.cormat() function is a list containing the following components :
      - r   : The table of correlation coefficients
      - p   : Table of p-values corresponding to the significance levels of the correlations
      - sym : A representation of the correlation matrix in which coefficients are replaced by symbols
              according to the strength of the dependence.

    In the generated graph,
        negative correlations are in blue and positive ones in red color.

(@) ###Upper triangle of the correlation matrix

#
    Note that in the result above, 
        only the lower triangle of the correlation matrix is shown by default. 
    
    You can use the following R script to get the upper triangle or the full correlation matrix.
    
```{r code5, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
rquery.cormat(mydata, type="upper")
```


(@) ###Full correlation matrix

```{r code6, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
rquery.cormat(mydata, type="full")
```


(@) ###Change the colors of the correlogram

```{r code7, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
col <- colorRampPalette(c("blue", "white", "red"))(20)
cormat <- rquery.cormat(mydata, type="full", col=col)
```


(@) ###Draw a heatmap

```{r code8, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
cormat<-rquery.cormat(mydata, graphType="heatmap")
```

+ To calculate the correlation matrix without plotting the graph, 
  you can use the following R script :
  
```{r code9, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
rquery.cormat(mydata, graph=FALSE)
```


(@) ###Format the correlation table

  + The R code below can be used to format the correlation matrix into a table of four columns containing :
      - The names of rows/columns
      - The correlation coefficients
      - The p-values
  
  + For this end, use the argument : type=“flatten”  
  
```{r code10, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
rquery.cormat(mydata, type="flatten", graph=FALSE)
```
  

(@) ###Description of rquery.cormat function

