---
title: "Regression Analysis 4"
output: html_document
---

##■ Model Predictions, Confidence Intervals and Prediction Intervals

The **main goal** of **linear regression** is to **predict** an outcome value on the basis of one or multiple **predictor variables**.
---

---

+ We’ll describe **how to predict** outcome for **new** observations data.  
+ You will also learn how to display the **confidence intervals** and the **prediction intervals**.

---

###1. Build a linear regression

We start by building a simple linear regression model that **predicts** the **stopping distances** of cars on the basis of the **speed**.

```{r eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}

data(cars)

str(cars)
```

---

```{r eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}

library(dplyr)
sample_n(tbl = cars, size = 5, replace = FALSE)
```

---

```{r eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}

library(tidyverse)
ggplot(cars, aes(speed, dist) ) + geom_point() + stat_smooth()
```

---

```{r eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}

# 1. Build the model
( model <- lm(dist ~ speed, data = cars) )

# 2. interpret the model
summary(model)
```

The **linear model equation** can be written as follow: **`dist = -17.579 + 3.932*speed`**.

#
    Note that, the units of the variable speed and dist are respectively, mph and ft.

---

###2. Prediction for new data set

Using the above model, we can **predict** the **stopping distance** for a **new speed** value.

Start by creating **a new data frame** containing, for example, **three new speed values** :

```{r eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}

( new.speeds <- data.frame(
  speed = sample(x = 1:100, size = 50, replace = FALSE)
) )
```

---

You can **predict** the corresponding **stopping distances** using the function **predict{stats}** as follow:

```{r eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}

predict(model, newdata = new.speeds)
```

---

###3. Confidence interval

The confidence interval reflects the **uncertainty around the mean predictions**. 

To display the **95% confidence intervals around the mean the predictions**,  
specify the option **interval = "confidence"** :

```{r eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}

predict(model, newdata = new.speeds, interval = "confidence")
```

The output contains the following columns:

    - fit         : The predicted stopping distance values for the new speed values
    - lwr and upr : The lower and the upper confidence limits for the expected values, respectively. 
                    By default the function produces the 95% confidence limits.

For example, the **95% confidence interval** associated with a **speed of 19 is (51.83, 62.44)**.

This means that, according to our model,  
A car with a **speed of 19 mph has, on average, a stopping distance ranging between 51.83 and 62.44 ft**.  
(**평균적으로 범위가 5.83 ~ 62.44 피트에 이르는** 정지거리를 가짐).

---

###4. Prediction interval

The **prediction interval** gives **uncertainty around a single value**. 

In the same way, as the confidence intervals, the **prediction intervals** can be computed as follow:

```{r eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}

predict(model, newdata = new.speeds, interval = "prediction")
```

The **95% prediction intervals** associated with a **speed of 19 is (25.76, 88.51)**.  
This means that, according to our model, **95% of the cars with a speed of 19 mph have a stopping distance between 25.76 and 88.51**.  
(**19mph 속도인 차량의 95%가, 25.76 ~ 88.51피트 사이에서, 어떤 정지거리를 가짐**)

#
    Note that, prediction interval relies strongly on the assumption that the residual errors are normally distributed with a constant variance. So, you should only use such intervals if you believe that the assumption is approximately met for the data at hand.

---

###5. Prediction interval or confidence interval?

A **prediction interval** reflects the **uncertainty around a single value**, while a **confidence interval** reflects the **uncertainty around the mean prediction values**. Thus, a **prediction interval** will be generally **much wider** than a **confidence interval** for the **same value**.

---

Which one should we use?  
The answer to this question **depends on** the **context** and the **purpose** of the analysis.  
Generally, we are **interested in specific individual predictions**, so a **prediction interval would be more appropriate**.  
Using a **confidence interval** when you **should** be using a **prediction interval** will greatly **underestimate** the **uncertainty** in **a given predicted value** (P. Bruce and Bruce 2017).

---

The R code below creates a **scatter plot** with:

  - The **regression line** in **blue**
  - The **confidence band** in **gray**
  - The **prediction band** in **red**
  
```{r eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}

# 0. Build linear model 
( model <- lm(dist ~ speed, data = cars) )

# 1. Add predictions 
( pred.int <- predict(model, interval = "prediction") )
( mydata <- cbind(cars, pred.int) )

# 2. Regression line + confidence intervals
library("ggplot2")
p <- ggplot(mydata, aes(speed, dist)) + geom_point() + stat_smooth(method = lm)

# 3. Add prediction intervals
p + geom_line(aes(y = lwr), color = "red", linetype = "dashed")+
    geom_line(aes(y = upr), color = "red", linetype = "dashed")
```

