---
title: "Regression Analysis 1"
output: html_document
---

### Ordinary Least Squares Regression ( OLS Regression)

---

0. Used data set
- women{datasets}
```{r code0, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}

str(women)
women
```

---

1. Simple linear regression
- $\hat{weight} = -87.52 + 3.45*height$
```{r code8.1, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}

( fit <- lm(formula = weight ~ height, data = women) )
summary(fit)

women$weight

fitted(fit)

residuals(fit)

plot(x = women$height, y = women$weight, xlab = 'Height (inches)', ylab = 'Weight (lbs)')
abline(fit)

```

---

2. Polynomial regression
- $\hat{weight} = 261.87818 - 7.34832 * height + 0.08306 * height^{2}$
```{r code8.2, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}

( fit2 <- lm(formula = weight ~ height + I(height^2), data = women) )
summary(fit2)

with(data = women, expr = {
  plot(x = height, y = weight, xlab = 'Height (inches)', ylab = 'Weight (lbs)')
  lines(x = height, y = fitted(fit2))
})
```

---

```{r code8.2.1, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}

( fit3 <- lm(formula = weight ~ height + I(height^2) + I(height^3), data = women) )
summary(fit3)


library(car)
scatterplot(
  weight ~ height, 
  data = women, 
  spread = FALSE, 
  smoother.args = list(lty=2), 
  pch = 19,
  main = 'Women Age 30 ~ 39',
  xlab = 'Height (inches)',
  ylab = 'Weight (lbs)' 
)

```

---

3. Used data set
- state.x77{datasets}
```{r eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}

class(state.x77)
dim(state.x77)

library(DT)
( datatable(state.x77) )

states <- as.data.frame(state.x77[, c('Murder','Population','Illiteracy','Income','Frost')])

library(dplyr)
sample_n(tbl = states, size = 10, replace = FALSE)
```

---

4. Correlation analysis between two variables in states data frame
```{r code8.3, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}

cor( states )

library(car)
scatterplotMatrix(
  x = states,
  spread = FALSE,
  smoother.args = list(lty = 2),
  main = 'Scatter Plot Matrix'
)

```

---

5. Multiple linear regression
```{r code8.4, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}

( fit <- lm(formula = Murder ~ Population + Illiteracy + Income + Frost, data = states) )
summary(fit)
```

---

6. Multiple linear regression with significant interaction between indepedent variables

- Used dataset : mtcars{datasets}
```{r code8.4.1, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}

class(mtcars)
dim(mtcars)

library(DT)
( datatable(mtcars) )
```

---

- Regression model : $\hat{mpg} = 49.81 - 0.12*hp - 8.22*wt + 0.03*hp*wt$
```{r code8.4.2, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
#( fit <- lm(formula = mpg ~ hp + wt + hp:wt, data = mtcars) )
( fit <- lm(formula = mpg ~ hp + wt + wt:hp, data = mtcars) )
summary(fit)
```

---

- Interaction Plot of hp*wt
```{r code8.4.3, eval=TRUE, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}

library(effects)
#( eft <- effect('hp:wt', fit,, list(wt = c(2,2, 3.2, 4.2))) )
( eft <- effect('hp:wt', fit,, list(hp = c(123.0, 146.7, 180))) )

class(eft)
str(eft)

plot(x = eft, multiline = TRUE)
```

---

